Architectures for workflow integration in science gateways

### 2016-04-12 paper.tex
Updates in paper.tex

\usepackage{svg}
\usepackage{url}
\usepackage{xspace}
\usepackage{enumitem}
\usepackage{hyperref}

@@

\begin{document}
\begin{frontmatter}

@@

@@

\item Workflow integration: consists in adding a new workflow to the
  science gateway so that users can execute it. It is conducted by
  science gateway administrators, developers or users, and it results
  in an interface, for instance a web form, where users can enter the
  parameters of the workflow to be executed. Integrating a workflow is
  not the same process as integrating a workflow engine.
\item Task control: operations to manage tasks on the infrastructure,
  including: submission, monitoring, termination, deletion, etc.
\item Data control: operations to manage data on
  the infrastructure, such as: upload, download, deletion, browsing,
  replication, caching, etc. Data movements can be triggered by the
  user, to upload input data or download processed data - interaction
  \texttt{(c1)}, or by the workflow engine, to transfer data across the
  infrastructure - interaction \texttt{(c2)}. \todo{add c1 and c2 to the
    Figures}
\item Workflow control: operations to execute a workflow, including:
  submission, monitoring, termination, etc. Workflow control can be
  coarse-grained (a.k.a. black box) or fine-grained (white box). In a
  coarse-grained model, workflow activities are masked. 
\item Sub-task control: operations used by tasks to submit sub-tasks
  on the infrastructure, including: submission, monitoring,
  termination, deletion, etc.
\item Pool-agent: specific to the architecture described in~\ref{sec:pool}.
\item Workflow conversion: translation from one workflow language to
  another one.
\end{enumerate}

@@

Complexity of the architecture (therefore development effort, robustness, maintainability, etc):
  * On the SG side
  * On the workflow engine side (what is required to wrap it).
Could be measured by counting the number of required functions,
e.g. submit job, call a workflow service, etc. Ponderated by the
"complexity" of the features between 1 (easy) and 3 (tricky). Look at
other types of measures.

Provided features:
  * Fine-grained workflow monitoring (debugging, progress report)
  * Scalability, load-balancing between engines
  * Support for multiple engines in the same gateway. Allows to reuse several different workflows in a single gateway but less complex than meta workflows. But do we really need meta workflows?
  * Support for meta workflows

"A Formal Approach to Support Interoperability in Scientific
Meta-workflows" (reviewed for IWSG and JoGC) has a formal model to
evaluate CGI and FGI.

"four major approaches for workflow interoperability include a,b,c,d" (see Terstyanzky et al, "Enabling scientific workflow sharing through CGI...", FGCS 2014) -> read this ref.

http://www.oreilly.com/programming/free/files/software-architecture-patterns.pdf

@@

\paragraph{Service}
+ Easy to implement in the SG (just call the service)
- Load-balancing between services is difficult
(refer to VIP's experience).
- Need to wrap workflow engines in
services.
- Need to know where to submit each workflow, e.g. when
multiple workflows are involved.
- The science gateway has no
knowledge of the detailed progress of the workflow, in particular
job-level information (statuses, stdout, stderr). If we wanted to do
that, specific feedback channels have to be developed, which requires more work (see VIP).

@@

@@

>>>>

\usepackage[x11names,dvipsnames,table]{xcolor} %for use in color links
\usepackage{svg}
\usepackage{url}
\usepackage{xspace}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{multirow}

@@

\begin{document}

\begin{frontmatter}

@@

\todo{Introduce the notion of architectural diagram with an abstract figure.}

@@

\item Workflow integration: consists in adding a new workflow to the
  science gateway so that users can execute it. It is conducted by
  science gateway administrators, developers or users, and it results
  in an interface, for instance a web form, where users can enter the
  parameters of the workflow to be executed. Integrating a workflow is
  not the same process as integrating a workflow engine.
\item Task control: operations to manage tasks on the infrastructure,
  including: submission, monitoring, termination, deletion, etc.
\item Data control: operations to manage data on
  the infrastructure, such as: upload, download, deletion, browsing,
  replication, caching, etc. Data movements can be triggered by the
  user, to upload input data or download processed data -- interaction
  \texttt{$c_1$}, or by the workflow engine, to transfer data across the
  infrastructure -- interaction \texttt{$c_2$}. \todo{add c1 and c2 to the
    Figures}
\item Workflow control: operations to execute a workflow, including:
  submission, monitoring, termination, etc. Workflow control can be
  coarse-grained (a.k.a. black box) or fine-grained (white box). In a
  coarse-grained model, workflow activities are masked. 
\item Sub-task control: operations used by tasks to submit sub-tasks
  on the infrastructure, including: submission, monitoring,
  termination, deletion, etc.
\item Pool-agent: specific to the architecture described in~\ref{sec:pool}.
\item Workflow conversion: translation from one workflow language to
  another one.
\end{enumerate}

@@

We use four main criteria to evaluate the architectures: integration
effort, robustness, modularity and scalability. Criteria break down to
specific metrics for which low values indicate good performance.

\todo{Explain these metrics better, including why the related features
  are useful.}

\emph{Integration effort} measures the development effort required to build
the system, assuming that none of the components could initially
communicate with each other. It is measured by counting the number of
interactions to develop and components to modify on the architectural
diagram. It breaks down to 2 metrics that relate to the direction of
interaction arrows on the architectural diagrams:
\begin{itemize}[itemsep=0cm]
\item Interactions from the science gateway. 
\item Iteractions from the workflow engine. 
\end{itemize}

\emph{Robustness} relates to the complexity of the architecture and
its ability to provide relevant debugging information about the
workflow executions. It breaks down to 3 metrics:
\begin{itemize}[itemsep=0cm]
\item Components: number of unique components involved in workflow executions.
\item Interactions: number of unique interactions involved in workflow executions. 
\item Debugging: availability of debugging information about workflow activities
  (1: difficult to obtain; 0: easy to access).
\end{itemize}

\emph{Modularity} measures the ability to replace elements in the
architecture, or to integrate new elements. Elements include
workflow engines, workflows and infrastructure. It breaks down in 4
metrics:
\begin{itemize}[itemsep=0cm]
\item New engine type: number of interactions or components to modify
  to integrate a new workflow engine type in the architecture.
\item New engine version: number of interactions or components to
  modify to integrate a new version of a workflow engine in the
  architecture, assuming that another version of the same engine type
  is already available.
\item New infrastructure: number of interactions or components to
  modify to integrate a new type of infrastructure in the
  architecture.
\item New workflow: number of interactions required to integrate a new
  workflow in the architecture.
\item Meta-workflow: ability to describe meta-workflows from existing
  workflows (1: not available; 0: available).
\end{itemize}

\emph{Scalability} measures the ability of the architecture to support
high workloads. It breaks down into 4 metrics:
\begin{itemize}[itemsep=0cm]
\item New engine instance (0: automated; 1: manual intervention needed).
\item Elastic engines (0: easy to implement; 1: difficult to implement).
\item Distributed engines (0: possible; 1: not possible).
\item Task scheduling (0: difficult; 1: more difficult).
\end{itemize}

%"A Formal Approach to Support Interoperability in Scientific
%Meta-workflows" (reviewed for IWSG and JoGC) has a formal model to
%evaluate CGI and FGI.
%"four major approaches for workflow interoperability include a,b,c,d" (see Terstyanzky et al, "Enabli%ng scientific workflow sharing through CGI...", FGCS 2014) -> read this ref.
% http://www.oreilly.com/programming/free/files/software-architecture-patterns.pdf

@@

See Table~\ref{table:evaluation}.

We focus on the integration of workflow engines in science gateways,
i.e. we assume that the science gateway is already interfaced with the
infrastructure.

\paragraph{Service invocation} Integrating a workflow engine as a
service requires 3 interactions. First, the science gateway has to be
extended to call the engine interface (\texttt{d}), which might
require some adaptation on the workflow engine too, for instance
wrapping in a web service. In addition, the workflow engine has to be
interfaced with the infrastructure, which requires the two
interactions \texttt{b} and \texttt{$c_2$}.

+ Easy to implement in the SG (just call the service)
- Load-balancing between services is difficult
(refer to VIP's experience).
- Need to wrap workflow engines in
services.
- Need to know where to submit each workflow, e.g. when
multiple workflows are involved.
- The science gateway has no
knowledge of the detailed progress of the workflow, in particular
job-level information (statuses, stdout, stderr). If we wanted to do
that, specific feedback channels have to be developed, which requires more work (see VIP).

@@

\begin{table*}
\centering
\begin{tabular}{rcccccc}
\textbf{Metric}                      & \textbf{Tight}
                                     & \textbf{Service}
                                     & \textbf{Sub-task}
                                     & \textbf{Pool}
                                     & \textbf{Nested}
                                     & \textbf{Import} \\
\multicolumn{7}{c}{\cellcolor[HTML]{EEEEEE}\textbf{Integration effort}}\\
From science gateway                 & \multirow{2}{*}{3}
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99FF99}2  
                                     & \cellcolor[HTML]{99FF99}2
                                     & \cellcolor[HTML]{99FF99}2
                                     & \cellcolor[HTML]{99AA99}3 \\
From workflow engine                 &
                                     & \cellcolor[HTML]{99FF99}1 
                                     & \cellcolor[HTML]{99FF99}2
                                     & \cellcolor[HTML]{99AA99}2
                                     & \cellcolor[HTML]{99AA99}2
                                     & \cellcolor[HTML]{99FF99}1 \\
\textbf{Total}                       & \cellcolor[HTML]{99FF99}\textbf{3}
                                     & \cellcolor[HTML]{99FF99}\textbf{3}
                                     & \cellcolor[HTML]{99FF99}\textbf{3}
                                     & \cellcolor[HTML]{99AA99}\textbf{4}
                                     & \cellcolor[HTML]{99AA99}\textbf{4}
                                     & \cellcolor[HTML]{99AA99}\textbf{4} \\

\multicolumn{7}{c}{\cellcolor[HTML]{EEEEEE}\textbf{Robustness}}\\
Components                           & \cellcolor[HTML]{99FF99}2
                                     & \cellcolor[HTML]{99DD99}3
                                     & \cellcolor[HTML]{99FF99}2
                                     & \cellcolor[HTML]{99AA99}4  
                                     & \cellcolor[HTML]{99AA99}4 
                                     & \cellcolor[HTML]{99AA99}4 \\
Interactions                         & \cellcolor[HTML]{99FF99}2    
                                     & \cellcolor[HTML]{99DD99}3 
                                     & \cellcolor[HTML]{99DD99}3      
                                     & \cellcolor[HTML]{99AA99}4  
                                     & \cellcolor[HTML]{99AA99}4      
                                     & \cellcolor[HTML]{99AA99}4 \\
Debugging                            & \cellcolor[HTML]{99AA99}0          
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99AA99}1
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99AA99}1
                                     & \cellcolor[HTML]{99FF99}0 \\
\textbf{Total}                       & \cellcolor[HTML]{99FF99}\textbf{4}
                                     & \cellcolor[HTML]{99CC99}\textbf{6}
                                     & \cellcolor[HTML]{99CC99}\textbf{6}
                                     & \cellcolor[HTML]{99BB99}\textbf{8}
                                     & \cellcolor[HTML]{99AA99}\textbf{9}
                                     & \cellcolor[HTML]{99BB99}\textbf{8} \\
\multicolumn{7}{c}{\cellcolor[HTML]{EEEEEE}\textbf{Modularity}}\\
New engine type                      & \cellcolor[HTML]{99AA99}3
                                     & \cellcolor[HTML]{99DD99}2    
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99DD99}2
                                     & \cellcolor[HTML]{99DD99}2
                                     & \cellcolor[HTML]{99AA99}3 \\
New engine version                   & \cellcolor[HTML]{99AA99}1
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0 \\
New infrastructure                   & \cellcolor[HTML]{99FF99}2
                                     & \cellcolor[HTML]{99FF99}2
                                     & \cellcolor[HTML]{99FF99}2
                                     & \cellcolor[HTML]{99FF99}2
                                     & \cellcolor[HTML]{99FF99}2
                                     & \cellcolor[HTML]{99FF99}2 \\
New workflow                         & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99AA99}2 \\
Meta-workflow                        & \cellcolor[HTML]{99AA99}1
                                     & \cellcolor[HTML]{99AA99}1  
                                     & \cellcolor[HTML]{99AA99}1
                                     & \cellcolor[HTML]{99AA99}1
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99AA99}1 \\
\textbf{Total}                       & \cellcolor[HTML]{99AA99}\textbf{8}
                                     & \cellcolor[HTML]{99EE99}\textbf{6}
                                     & \cellcolor[HTML]{99FF99}\textbf{5}
                                     & \cellcolor[HTML]{99EE99}\textbf{6}
                                     & \cellcolor[HTML]{99FF99}\textbf{5}
                                     & \cellcolor[HTML]{99AA99}\textbf{8} \\
\multicolumn{7}{c}{\cellcolor[HTML]{EEEEEE}\textbf{Scalability}}\\
New engine instance                  & \cellcolor[HTML]{99AA99}1
                                     & \cellcolor[HTML]{99AA99}1
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99AA99}1
                                     & \cellcolor[HTML]{99AA99}1 \\
Elastic engines                      & \cellcolor[HTML]{99AA99}2
                                     & \cellcolor[HTML]{99DD99}1
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99DD99}1
                                     & \cellcolor[HTML]{99DD99}1 \\
Distributed engines                  & \cellcolor[HTML]{99AA99}1
                                     & \cellcolor[HTML]{99AA99}1
                                     & \cellcolor[HTML]{99AA99}1
                                     & \cellcolor[HTML]{99AA99}1
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99AA99}1 \\
Task scheduling                      & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99AA99}1
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0 \\
\textbf{Total}                       & \cellcolor[HTML]{99AA99}\textbf{4}
                                     & \cellcolor[HTML]{99BB99}\textbf{3}
                                     & \cellcolor[HTML]{99DD99}\textbf{2}
                                     & \cellcolor[HTML]{99FF99}\textbf{1}
                                     & \cellcolor[HTML]{99DD99}\textbf{2}
                                     & \cellcolor[HTML]{99BB99}\textbf{3} \\
\multicolumn{7}{c}{\cellcolor[HTML]{EEEEEE}\textbf{Grand total}}\\
                                     & \cellcolor[HTML]{99CC99}\textbf{19}
                                     & \cellcolor[HTML]{99BB99}\textbf{18}
                                     & \cellcolor[HTML]{99FF99}\textbf{16}
                                     & \cellcolor[HTML]{99CC99}\textbf{19}
                                     & \cellcolor[HTML]{99DD99}\textbf{20}
                                     & \cellcolor[HTML]{99AA99}\textbf{23}
\end{tabular}
\caption{Architecture evaluation. Lower values (brighter colors) indicate better performance.}
\label{table:evaluation}
\end{table*}

@@

\section*{References}
